---
title: "Computational Statistics Lab01"
author: "Tejashree R Mastamardi"
date: "7/7/2020"
output: 
  pdf_document
  toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(dplyr)
library(ggplot2)
```

#The exponential
```{r message=FALSE, warning=FALSE}
#Taylor Series
Taylor <- function(x, N){
  1+sum(sapply(1:N, function(i,x){x^i/prod(1:i)},x=x,simplify=TRUE))
}
exp(20)
Taylor(20,100)
Taylor(20,100) - exp(20)
exp(-20)
Taylor(-20,100)
Taylor(-20,200)
```

```{r message=FALSE, warning=FALSE}
f1 <- function(x){
  (x-1)^6
}
f2 <- function(x){
  1 - 6*x + 15*x^2 - 20*x^3 + 15*x^4 - 6*x^5 + x^6
}
x <- seq(from=0.995, to=1.005, by=0.0001)
y1 <- f1(x)
y2 <- f2(x)

plot(x, y1, pch = 19, cex = 0.5, ylim = c(-5*10^(-15), 20*10^(-15)), main = "Two ways to calculate(x-1)^6", xlab = "x", ylab = "y")
points(x, y2, pch = 18, cex = 0.8)
```

Question 1: Be careful when comparing
1.1
```{r message=FALSE, warning=FALSE}
x1 <- 1/3 
x2 <- 1/4
if ( x1 - x2 == 1/12) {
print ( " Subtraction is correct " )
} else {
print ( " Subtraction is wrong" )
}
```

```{r message=FALSE, warning=FALSE}
x1 <- 1
x2 <- 1/2
if ( x1 - x2 == 1/2) {
print ( " Subtraction is correct " )
} else {
print ( " Subtraction is wrong" )
}
```

1.2
#Improvement
```{r message=FALSE, warning=FALSE}
x1 <- 1/3
x2 <- 1/4
isTRUE(all.equal(x1-x2,1/12))
```

1.2
#Improvement
```{r message=FALSE, warning=FALSE}
x1 <- 1/3 
x2 <- 1/4
if(all.equal(x1 - x2, 1/12)){
print ( " Subtraction is correct " )
} else {
print ( " Subtraction is wrong" )
}
```

Question 2: Derivative
2.1
```{r message=FALSE, warning=FALSE}

f_x <- function(x){
  e <- 10^-15
  ((x + e) - x)/e
}
```

2.1 alternative method
```{r message=FALSE, warning=FALSE}
e <- 10^-15
f_x <- function(x){
  ((x+e)-x)/e
}
```

2.2
```{r message=FALSE, warning=FALSE}
#at x=1
f_x(1)

#at x=100000
f_x(100000)
```

2.3
We can see that when adding the very small number e = 10−15 to x = 100000, the result stays
100000. We have an underflow problem since decimals are lost. When adding the same small
number e = 10−15 to x = 1, the underflow does not occur because x is a relatively small number
and decimals can be saved as well up to some point.


For x=1, Value obtained is 1.110223 and for x=100000 Value obtained is 0. The true values are
1. When x=1, adding epsilon = 10ˆ-15 which increases the magnitude of x by a very small amount. When
x=10000, adding epsilon = 10ˆ-15 does not make much of a difference to the magnitude of x because of which
the numerator becomes 0(zero) leading the whole equation to become 0(zero). We have an underflow problem because of which the decimals are lost.

Question 3: Variance
3.1
```{r message=FALSE, warning=FALSE}
MyVar <- function(x){
  n = length(x)
  result = (1/(n-1))*(sum(x^2) - ((1/n)*(sum(x)^2)))
  return(result)
}

myvar <- function(x){
  n = length(x)
  result = (1/n-1) * (sum(x^2)-((1/n)*(sum(x)^2)))
  return(result)
}
```

3.2 Generate a vector with 10000 random numbers with mean 10^8 and variance 1.
```{r message=FALSE, warning=FALSE}
x <- rnorm(10000, mean = 10^8, sd = 1)
plot(x)
```

3.3 Compute the difference Yi=myvar(xi)-var(xi)
```{r message=FALSE, warning=FALSE}
Yi<-c()
for (i in 1:length(x)) {
Yi[i]<-MyVar(x[1:i])-var(x[1:i])
}
plot(x=1:length(Yi),y=Yi)
plot(Yi[1:i])#either of the two plot chunks as both plots are same
```

3.4 
```{r message=FALSE, warning=FALSE}
#calculate mu values
mu<-function(x){
(1/length(x)*sum(x))
}

#calculate the variance values
myvar2<-function(x){
1/length(x)*sum((x-mu(x))^2)
}

#variance estimator minus built in function var() in R
Yi2<-c()
for (i in 1:length(x)) {
Yi2[i]<-myvar2(x[1:i])-var(x[1:i])
}

#plot
plot(x=1:length(Yi2),y=Yi2)
plot(Yi2[1:i])#either of the two plot chunks as both plots are same
```

Question 4: Linear Algebra
4.1
```{r message=FALSE, warning=FALSE}
tecator <- read.csv("tecator.csv")
tecator
```
4.2
```{r message=FALSE, warning=FALSE}
x <- as.matrix(tecator[,c(2:102,104)])
y <- tecator[,c(103)]

xt <- t(x)
a <- xt %*% x #use modulus for multiplying matrices in R
b <- xt * y

#matrix X are the observations of the absorbance records,levels of moisture and fat
X<-tecator[,c(2:102,104)]#X
X<-as.matrix(X)#Because we are multiplying matrix so we need to pass the data frame in to as.matrix

#y are the proteinlevels
y<-tecator[,c(103)]#Y is a vector with one coloum

#Transpose of X=XT
XT<-t(X)

#A = XTX
A <- XT%*%X 

#b = XTy
b<-XT%*%y
```


4.3
```{r message=FALSE, warning=FALSE, echo=FALSE}
solve(a,b)
```
we got the result as,“Error in solve.default(A,b):system is computationally singular reciprocal condition
number = 7.13971e-17”

4.4
```{r message=FALSE, warning=FALSE}
condition_number<-kappa(a)
condition_number
```

```{r message=FALSE, warning=FALSE}
#scaling the data 
#2step-matrix X are the observations of the absorbance records,levels of moisture and fat
X<-tecator[,c(2:102,104)]#X scaled
X_s<-scale(X)
X_s<-as.matrix(X_s)

# y are the scaled protein levels
y<-tecator[,c(103)]#Y is a vector with one column
y_s<-scale(y)

#Transpose XT-scaled 
XT_s<-t(X_s)

#A = XTX
A_s <- XT_s%*%X_s

#b = XTy
b_s<- XT_s%*%y_s

#3
solve(A_s,b_s)

#4step kappa()with scaled data
condition_number_s<-kappa(A_s)# condition number is the //A// //A transpose//
condition_number_s
```
The kappa value obtained for the scaled matrix is small compared to the kappa value obtained by the original matrix(without scaling) which seems to be a good sign and because of this small kappa value computations were possible.
