---
title: "Lab1-Computational Stastics"
author: "Prudhvi Peddmallu (prupe690), Tejashree R Mastamardi (tejma768)"
date: "21 January 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Question 1

```{r message=FALSE, warning=FALSE}
x1<-1/3
x2<-1/4
if (x1-x2==1/12){
  print("Subtraction is correct")
}else{
  print("Subtraction is wrong")
}

```

Explanation:- We are getting the output as “subtraction is wrong”, because the number of digits after the decimals for x1-x2 and 1/12 do not the match.

```{r message=FALSE, warning=FALSE}

x1<-1
x2<-1/2
if (x1-x2==1/2){
  print("Subtraction is correct")
}else{
  print("Subtraction is wrong")
}
```

Explanation:- We are getting the output as “subtraction is correct”, because the number of digits after the decimals for x1-x2 and 1/2 are the same.

```{r message=FALSE, warning=FALSE}

x1 <- 1/3
x2 <- 1/4
if(all.equal((x1-x2),(1/12))){
print("subtraction is correct")
}else{
print("subtraction is wrong")
}

```

#Question 2: Derivative


```{r message=FALSE, warning=FALSE}

#2.1
Derivative <- function(x,e) {
a <- ((x+e)-x)/e
return(a)
}
x <-1
e<-10^-15
Derivative(x,e)

#2.2
x <- 10000
e<-10^-15
Derivative(x,e)

```
 For x=1, Value obtained is 1.110223 and for x=100000 Value obtained is 0. 
 The true values are 1. 
 When x=1, adding epsilon which is 10^-15 which increases the magnitude of x by a very small amount. 
 When x=10000, adding epsilon which is 10^-15 does not make much of a difference to the magnitude of x because of which the numerator becomes
0 leading the whole equation to become 0.


#Question 3 Variance

```{r message=FALSE, warning=FALSE}

set.seed(12345)
x<-rnorm(10000,mean= 10^8,sd=1)#10000 is the sequence of observations i to 1 to 100000

myvar<-function(x){
    1/(length(x)-1)*((sum(x^2)- 1/length(x)*(sum(x))^2))# this is the formula of variance here n is the length of x observations 
  }
Yi<-c()
for (i in 1:length(x)) {
     Yi[i]<-myvar(x[1:i])-var(x[1:i])
     }
plot(x=1:length(Yi),y=Yi)

```

#Question 4: Linear Algebra

```{r message=FALSE, warning=FALSE}

#1
library(dplyr)
library(ggplot2)
#Reading the file tecator in to R
tecator<-read.csv("tecator.csv")

```

```{r message=FALSE, warning=FALSE}

#2
#matrix X are the observations of the absorbance records,levels of moisture and fat
X<-tecator[,c(2:102,104)]#X
X<-as.matrix(X)#Because we are multiplying matrix so we need to pass the data frame in to as.matrix

#y are the proteinlevels
y<-tecator[,c(103)]#Y is a vector with one coloum
#Transpose XT
XT<-t(X)
#A = XTX
A <- XT%*%X 

#b = XTy
b<-XT%*%y   

```

```{r message=FALSE, warning=FALSE}
#3 solve()# it will show the error 
solve(A,b) 
#Error in solve.default(A) : system is computationally singular: reciprocal condition number = 7.13971e-17

```


```{r message=FALSE, warning=FALSE}
#4 kappa()
#kappa Compute or Estimate the Condition Number of a Matrix 
condition_number<-kappa(A)
condition_number
```

The function kappa () compute or estimate the condition  number of a matrix .The condition number of A is A mutlipled  with A transpose. and this condition number(kappa()) depends on the norm.If the condition number of a matrix is larger value then it is a"bad" condition because several orders of a magnitude.

```{r message=FALSE, warning=FALSE}
#5
#scaleing the data 
# 2step-matrix X are the observations of the absorbance records,levels of moisture and fat
X<-tecator[,c(2:102,104)]#X scaled
X_s<-scale(X)
X_s<-as.matrix(X_s)
# y are the proteinlevels -scaled
y<-tecator[,c(103)]#Y is a vector with one coloum
y_s<-scale(y)
#Transpose XT-scaled 
XT_s<-t(X_s)
XT_s
#A = XTX
A_s <- XT_s%*%X_s
#b = XTy
b_s<- XT_s%*%y_s
#3
solve(A_s,b_s)
#4step kappa()with scaled data
condition_number_s<-kappa(A_s)# condition number is the //A// //Atranspose//
condition_number_s
```
 


