---
title: "Lab3"
author: "Prudhvi Peddmallu (prupe690), Tejashree R Mastamardi (tejma768)"
date: "February 11, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE, message = FALSE, warning = FALSE,	comment = NA)
```


#Question1:Cluster Sampling

```{r}
#1 Importing file to R.

my_data <- read.csv2("population.csv")

```


```{r}
#2 Use a uniform random number generator to create a function that 
#selects 1 city from the whole list by the probability scheme offered above.

Select_City <- function(City){
probability <- City$Population/sum(City$Population)
cumulativesumprob <- cumsum(probability)
i <- runif(1, min = 0, max = 1)
return(which.max(i < cumulativesumprob))
}

```


```{r}
#3 Using the function created in Step 2 to give exactly 20 cities.

city_pop <- data.frame()
for (i in 1:20){
Index_city <- Select_City(my_data)
city_pop[i,1] <- my_data$Municipality[Index_city]
city_pop[i,2] <- my_data$Population[Index_city]
my_data <- my_data[-Index_city,]
}
colnames(city_pop) <- c("City","Population")

```

```{r}
#4 Selected 20 cities.

city_pop

```
 
 Explanation:-
 Selected 20 cities are listed above, The size(population) of the selected 20 cities is quite large in number.
 
```{r}
#5 Plot of histogram showing size of all cities in the Country
histogram1 <- hist(my_data$Population, col = "red", breaks = 50, xlab = "Population", 
                 ylab = "All of cities")

```
```{r}
#Plot of histogram showing the size of 20 selected cities
histogram2 <- hist(city_pop$Population, col = "light blue", breaks = 50, xlab = "Population", 
                   ylab = "Number of cities is 20")

```
 
 From the two histogram plots it is evident that they follow a similar distribution. This means that the function that we have implemented to select 20 cities seems to be doing a good job in the selection process.
 

#Question 2: Different distributions
The double exponential (Laplace) distribution is given by formula:

$$DE(\mu,\alpha) = \frac{\alpha}{2} e^{(-\alpha|x-\mu|)}$$

The Cumulative Distribution Function is given by 

$$F(x)=\int_{-\infty}^{x} f(x) dx$$

$$F(x)=\int_{-\infty}^{x} \frac{\alpha}{2}e^{-\alpha(x-\mu)} dx ,~~~~~(if~~x>\mu)$$

$$F(x)=1-\int_{x}^{\infty} \frac{\alpha}{2}e^{-\alpha(x-\mu)} dx$$

$$F(x)=1-\frac{1}{2}e^{-\alpha(x-\mu)}$$

$$F(x)=\int_{-\infty}^{x} \frac{\alpha}{2}e^{\alpha(x-\mu)} dx ,~~~~~if~~x\leq\mu$$

$$F(x)=\frac{1}{2}e^{\alpha(x-\mu)}$$

Inverse of CDF

$$For~~x>\mu,~we~got~F(x)=1-\frac{1}{2}e^{-\alpha(x-\mu)}$$

$$y=1-\frac{1}{2}e^{-\alpha(x-\mu)}$$

$$\frac{ln(2-2y)-\alpha\mu}{-\alpha}=x$$
$$For~U\sim U(0,1),~~~~~\frac{ln(2-2U)-\alpha\mu}{-\alpha}=X$$

$$For~~x\leq\mu,~we~got~F(x)=\frac{1}{2}e^{\alpha(x-\mu)}$$
$$y=\frac{1}{2}e^{\alpha(x-\mu)}$$

$$\frac{ln(2y)}{\alpha}+\mu=x$$

$$For~U\sim U(0,1),~~~~~\frac{ln(2U)}{\alpha}+\mu=X$$


```{r}
#1 Write a code generating double exponential distribution DE(0; 1) from Unif(0; 1) by using the inverse CDF method.

DE <- function(mu, a){
  U <- runif(1,0,1)
  if (U > 0.5){
        distribution <- (log(2-2*U)-a*mu)/(-a)
      }
      else {
        distribution <- (log(2*U)/a)+mu
      }
  return(distribution)
}

values <- c()
for (i in 1:10000){
values[i] <- DE(0,1)
values
}

hist(values, breaks = 1000, main = "Inverse CDF Method used to Generate Random Numbers")

```


```{r}
#2 Use the Acceptance/rejection method with DE(0; 1) as a majorizing density to 
# generate N(0; 1) variables.

accept_reject <- function(){
X<-c()
counter= 0 # counter for how many numbers been rejected
while (length(X)==0) { # as long as our vector is empty >> do
counter = counter+1
Y<-DE(0,1)
fy <- (1/sqrt(2*pi))*exp(-(1/2)*Y^2)
Gy <- 0.5*exp(-abs(Y))
c <- sqrt(2*exp(1)/pi)
U = runif(1,0,1)
if (U <= fy/(c*Gy)){
X=Y
}}
gen_itr <- cbind("gen" = X, "itr" = counter)
return(gen_itr)
}

```


```{r}
# generating 2000 random numbers using Acceptance/Rejection method
res <- list()
for (i in 1:2000){
res[[i]]<- accept_reject()
}
```

```{r}
# excluding the random numbers from the list for plot
gen <- c()
for (i in 1:2000){
gen[i]<- res[[i]][1]
}
# plotting the generated random numbers using N(0,1)
hist(gen , breaks = 50, main = "Random numbers generated using Acceptance-Rejection Method")

```


```{r}
# getting the total number of iterations (accepted+rejected) to commpute the average
itr <- c()
for (i in 1:2000){
itr[i]<- res[[i]][2]
}
# average rejection rate
average_rej <-1-(2000/sum(itr))
average_rej

```

```{r}
# expected rejection rate
c <- sqrt(2*exp(1)/pi)
expected_rej_rate<- 1-(1/c)
expected_rej_rate
```

We have chosen the constant c as:-
 $$constant = 2*e^{(1/2)/sqrt(2*pi)}$$
The value of Average Rejection rate R = 0.24 and the value of Expected Rejection rate ER = 0.33. ER(Expected Rejection rate) is close to R(Average Rejection rate) by 0.1

 